Forecasting the black Sigatoka development rate:\\ A comparison of machine learning techniques

### 2016-06-18 techniques.tex
Editando las primeras 3 técnicas.  Unifiqué la notación pero hay que ver si se logra esa unificación en todo el paper

\section{Machine learning techniques and Related works}
\label{sec:techs}

In this section we summarize the machine learning techniques used and we present the main related works.

@@

From the perspective of Support Vector Regression (SVR) the regression
function $y = f(s)$ for a given dataset $D=\{(s_i,y_i)\}_{i=1}^n$ is
represented as a linear function of the form \citep{Wei2013}:
\begin{equation*}
  f(s)=w^Ts+b
\end{equation*}
where $w$ and $b$ are respectively the weight vector and the intercept
of the model, and they are selected to find an \todofn{optimal}%
{optimal respecto a qué?.  Hay muchísimas posibilidades de optimalidad!}
%
fit to the data available in $D$ \review{in the sense of ...?}

For nonlinear cases, the $p$-dimensional input vectors are mapped by a
nonlinear function $\phi : R^p\rightarrow F$ onto a feature space
$F$.  After the nonlinear mapping, the regression function evolves to a
pervasive form:
\begin{equation*}
f(s)=w^T \phi (s)+b
\end{equation*}

In SVR the generalization ability is measured by the Euclidean norm of
$w$, and the empirical error is measured by the $\epsilon$-insensitive
loss function, given by \citep{Cristiani2005}

\TODO{si $l$ es una función, es función de qué, de $s$? $l(s)$? o de
  $y$? o de todos $l(y,s;\epsilon)$?}

\begin{equation*}
l= {\left| y - f(s) \right| }_{\epsilon } = 
\begin{cases}
   0 & {\left| y - f(s) \right| } \leq \epsilon \\
   \left| y - f(s) \right| - \epsilon & \text{otherwise}
\end{cases}
\end{equation*}
which ignores the error if the difference between the prediction value
and the actual value is smaller than $\epsilon$.
%
The $\epsilon$-insensitive loss function allows to find the
coefficients $w$ and $b$ by solving a convex optimization problem,
which balances the empirical error and the generalization ability.
%
Then, the optimization problem to identify the regression model can be
formulated as \citep{Wei2013}:
%
\begin{equation} 
\begin{aligned}
& \underset{}{\text{minimize}}
& & J(w,\xi_i , \xi_i^* ) = \frac{1}{2}   \Bigr| \Bigr| w \Bigr| \Bigr|^2 + C \sum_{i=1}^{n} (\xi_i , \xi_i^* )    \\
& \text{subject to}
& & \begin{array}{lcl} 
y_i - w^T \phi (s) - b  \leq \epsilon + \xi_i   \\
w^T \phi (s) + b - y_i \leq \epsilon + \xi_i^* & i= 1,2,...,n \\
\xi_i , \xi_i^* \geq 0  \\
\end{array}
\end{aligned}
\end{equation}  
where $C$ denotes the penalty parameter between empirical and
generalization errors, and $\xi_i , \xi_i^*$ are slack
variables. \figref{figura2} shows this situation.

\begin{figure}[h] 
 \centering
 \includegraphics[scale=.9]{SVR}
 \caption{$\epsilon$-insensitive loss function \citep{Wei2013}.} 
 \label{figura2} 
\end{figure}

The solution of this optimization problem by the Lagrange method is given by:
\begin{equation*}
f(s) = w^T \phi (s) + b = \sum_{i=1}^{n} (\alpha_i - \alpha_i^*) K (s,s_i) + b
\end{equation*}
where $\alpha_i - \alpha_i^*$ are the Lagrange multipliers of the
optimization problem’s dual form and $K(s_i,s_j )$ is the kernel
function satisfying the Mercer condition, and holds:
\begin{equation*}
K(s_i,s_j ) = \big \langle  \phi(s_i) , \phi(s_j)  \big \rangle
\end{equation*}
Operations in the kernel function $K(s,s_i )$ are performed in the
input space rather than in the potentially high dimensional feature
space of $\phi$ \citep{Alonso2013}.

This method fits a linear model with coefficients $w = (w_1,..,w_p)$
to minimize the residual sum of squares between the observed responses
in the dataset, and the responses predicted by the linear
approximation. Mathematically it solves a problem of the form
\citep{scikitlearn2011}:
$$\min_{w} \Bigr| \Bigr| Xw - y \Bigr| \Bigr|_2^2  $$
where $X$ denotes the features matrix.

According Pedregosa et al. \citep{scikitlearn2011}  the coefficient estimates for Ordinary Least Squares rely on the independence of the model terms. When terms are correlated and the columns of the design matrix $X$ have an approximate linear dependence, the design matrix becomes close to singular and as a result, the least-squares estimate becomes highly sensitive to random errors in the observed response, producing a large variance. This situation of multicollinearity can arise, for example, when data are collected without an experimental design.

\subsection{Ridge regression}
The ridge regression addresses some of the problems of ordinary least squares regression by imposing a penalty on the size of the coefficients. The ridge coefficients minimize a penalized residual sum of squares  \citep{scikitlearn2011}:

$$\min_{w} { \Bigr| \Bigr| Xw - y \Bigr| \Bigr|_2^2  + \alpha \Bigr| \Bigr| w \Bigr| \Bigr|_2^2 } $$

Here, $\alpha \ \textgreater \  0$ is a complexity parameter that controls the amount of shrinkage: the larger the value of $\alpha$, the greater the amount of shrinkage and thus the coefficients become more robust to collinearity.

\subsection{Elastic-Net regression}
Elastic-Net is a linear regression model trained with $L1$ and $L2$ prior as regularizer. This combination allows for learning a sparse model where few of the weights are non-zero like Lasso, while still maintaining the regularization properties of Ridge  \citep{scikitlearn2011}. The convex combination of $L1$ and $L2$ is controled by using the $l1_{ratio}$ parameter.

>>>>

\section{Compared regression techniques}
\label{sec:techs}

In the prediction of the development rate of the black Sigatoka, we
compare techniques such as least squares or ridge regression, commonly
encountered in the agricultural literature with machine learning
methods such as support vector regression, elastic regression and echo
state networks, where the parameter space of each technique is also
taken into account.

Given a data set 
\begin{equation}
  D=\left\{(\vct{x}_i,y_i) \mid i=1\ldots n\right\}
\end{equation}
composed of the $d$-dimensional%
\footnote{Without loss of generality assume that the first component
  of every vector $\vct{x}_i$ is always $1$.}
%
feature vectors $\vct{x}_i\in\setR^d$ and the corresponding responses
$y_i$.
%
The ordinary least squares regression (OLSR) fits a linear model
$\tilde{y}_i = f(\vct{x}_i) = \vct{w}^T\vct{x}_i$ such that the sum of
squares of the residuals $(\tilde{y}_i-y_i)$ is minimized.
%
Let $\mat{X}$ be the $n\times{}d$ feature matrix containing the data
samples in its rows and $\vct{y}$ contain all the responses $y_i$
corresponding to each row, then the least squares regression finds
\begin{equation*}
  \hat{\vct{w}} =
  \arg\min_{\vct{w}}\left\| \mat{X}\vct{w} - \vct{y}\right\|_2^2
\end{equation*}
which is solved by means of the pseudoinverse
($\hat{\vct{w}}=\mat{X}^\dagger\vct{y}$) or equivalently by the singular
value decomposition of $\mat{X}$ \citep{Press2007}.

\TODO{Alex, no sé hasta donde hiciste en las pruebas ese ``arreglo''
  de que los datos tengan todos una entrada igual a 1.  De no ser así,
  el método no puede encontrar el offset, y sería como forzar que el
  modelo tenga que pasar por cero...}

\subsection{Ridge regression}

In contrast to the OLSR, the ridge regression (RR) adds a term to
penalize large weights \citep{scikitlearn2011}:
\begin{equation*}
\hat{\vct{w}} = 
\arg\min_{\vct{w}} \left\| \mat{X}\vct{w} - \vct{y} \right\|_2^2 + 
                 \alpha \left\|\vct{w}\right\|_2^2 
\end{equation*}
where the parameter $\alpha>0$ controls how strong is the shrinking of
the estimates towards zero.  This shrinkage introduces some bias but
helps to reduce the variance of the estimate.
%
The solution is given by $\hat{\vct{w}}=(\mat{X}^T\mat{X} +
\lambda\mat{I})^{-1}\mat{X}^T \vct{y}$.

@@

From the perspective of Support Vector Regression (SVR) the regression
function is usually stated as 
\begin{equation}
\label{eq:svrfunc}
  \tilde{y} = f(\vct{x}) = \vct{w}^T\vct{x} + b
\end{equation}
The weights are selected in a convex optimization problem
\citep{Smola2004}:
\begin{align*}
  \text{minimize}   & \frac{1}{2}\|\vct{w}\|^2 + C\sum_{i=1}^n(\xi_i+\xi_i^\ast \\
  \text{subject to} & 
  \begin{cases}
    y_i - \vct{w}^T\vct{x}_i - b &\leq \epsilon + \xi_i    \\
    \vct{w}^T\vct{x}_i + b -y_i  &\leq \epsilon + \xi_i^\ast\\
    \xi_i,\xi_i^\ast               &\geq 0
  \end{cases}
\end{align*}
where $\epsilon$ is the maximal allowed deviation of the targets
$\tilde{y}_i$ from $y_i$, the slack variables $\xi_i$ and $\xi_i^\ast$
allow to cope with otherwise unfeasible constraints for the
optimization problem, and the constant $C>0$ controls the trade-off
between the flatness of $f$ and the tolerance to deviations larger
than $\epsilon$.

Note that since OLSR and RR use a squared error function, data
outliers will have a strong influence on the resulting weights
$\vct{w}$.  On the SVR formulation, however, the usage of the $L_1$
norm and the slack variables considerably restrict or completely block
the influence of those outliers.

The SVR problem is reformulated by means of the dual optimization
problem into
\begin{equation}
  \label{eq:svrw}
  \vct{w}=\sum_{i=1}^n(\alpha_i-\alpha_i^\ast)\vct{x}_i
\end{equation}
where $\alpha_i,\alpha_i^\ast\in[0,C]$ are Lagrange multipliers
subject to $\sum_{i=1}^n(\alpha_i-\alpha_i^\ast)=0$.
%
In this so-called \emph{Support Vector expansion} the weights are
expressed as a linear combination of the data set patterns
$\vct{x}_i$.
%
Inserting (\ref{eq:svrw}) in (\ref{eq:svrfunc}) leads to
\begin{equation}
  \label{eq:svrf}
 f(\vct{x}) = \sum_{i=1}^n(\alpha_i-\alpha_i^\ast)\vct{x}_i^T\vct{x} + b
\end{equation}
The Lagrange multipliers $\alpha_i,\alpha_i^\ast$ are both non-zero
only for those data points where $|f(\vct{x}_i) - y_i|\geq\epsilon$.
Hence, the expansion of $\vct{w}$ in terms of $\vct{x}_i$ is sparse.
Those data points with non-vanishing coefficients are called
\emph{Support Vectors} \citep{Wei2013}.

Additionally, in (\ref{eq:svrf}) it is possible to employ the
\emph{kernel trick} and replace the terms $\vct{x}_i^T\vct{x}$ with
the evaluation of any Mercer kernel
$k(\vct{x}_i,\vct{x})=\phi(\vct{x}_i)^T\phi(\vct{x})$, where
$\phi(\vct{x})$ is a non-linear mapping of the input space onto a
higher (even infinite) dimensional feature space.
%
The kernel evaluation draws unnecessary the explicit evaluation of the
non-linear mapping.
%
This allows to solve non-linear regressions in the input space by
implicitly mapping through the kernel the samples into the higher
dimensional space, where the linear regression occurs \citep{Alonso2013}.

\TODO{Habría que insertar aquí los kernels utilizados en los experimentos}

\subsection{Elastic-Net regression}

Elastic-Net is a linear regression model trained with $L1$ and $L2$ prior as regularizer. This combination allows for learning a sparse model where few of the weights are non-zero like Lasso, while still maintaining the regularization properties of Ridge  \citep{scikitlearn2011}. The convex combination of $L1$ and $L2$ is controled by using the $l1_{ratio}$ parameter.
